Composite Bayesian inference

### 2018-11-16 notes/composite_inference.tex
Modifs 16/11/18

Again, what is the matter? The {\em real} difference between generative and discriminative models is that the former have all their parameters describing $p(y|x)$ while the latter have all their parameters describing $p(x|y)$. It's a question of parameters at the end of the day.

Therefore, it's a question of learning. We can argue that generative models can ``learn faster'' than discriminative ones and it's because the learning process involves ``understanding the data''. When you model data by an empirical distribution, you clearly don't understand a shit about the data. You can make fairly good predictions but you're like a monkey.

>>>>

Again, what is the matter? The {\em real} difference between generative and discriminative models is that the former have all their parameters describing $p(y|x)$ while the latter have all their parameters describing $p(x|y)$. Is it a question of parameters at the end of the day?

No, it's a question of what the models describe. A generative model describes the process of acquiring data, it has to reflect some physical understanding. A discriminative model describes a convenient approximating family for the recovery of a state, there is no physics involved. As a by-product, the generative model might well produce a good predictive distribution for the data, but it does that without understanding anything about the data.

Just like an histogram is a pretty dumb but good predictor if we have a shit load of data to compute it. Instead, you can have a parametric model that reflects some understanding of what the distribution is. It takes much less data.

La vraie question, c'est de savoir quel est le lien de cause \`a effet entre $x$ et $y$. Si $x$ est la ``cause'' de $y$, alors le mod\`ele g\'en\'eratif est naturel. Si c'est l'inverse, alors le mod\`ele discriminant est une \'evidence.

Par exemple, supposons que $y$ soit une vitesse moyenne et $x$ la distance parcourue en un temps donn\'e. On a $x=t y$ et c'est m\^eme d\'eterministe, donc aucune question \`a se poser: ce mod\`ele purement discriminant dans ce contexte donne la r\'eponse.

On peut se dire que la vitesse est la cause de la distance parcourue, mais la vitesse se d\'eduit tout aussi bien de la distance: $y=x/t$. Donc on pourrait aussi s'en sortir avec un mod\`ele g\'en\'eratif.

Si maintenant notre mesure $y$ est bruit\'ee, il n'y a plus d'ambigu\"it\'e: on \'ecrit $y=x/t+\epsilon$, c'est-\`a-dire qu'on travaille avec un mod\`ele g\'en\'eratif. La vitesse mesur\'ee devient la ``cons\'equence'' de la distance \`a cause du bruit de mesure.

On voit bien que la notion de causalit\'e est assez foireuse.

On voit aussi que la d\'efinition officielle (Wikipedia et cie) de l'inf\'erence statistique est tr\`es restrictive: tirer des conclusions sur une population \`a partir d'un \'echantillon. C'est minable. Mais la d\'efinition permet de se mettre dans le cadre rassurant de l'\'echantillonnage. L'\'echantillon $y$ est bien ``g\'en\'er\'e'' \`a partir de la population.

Je pr\'ef\`ere la d\'efinition de Floyd Bullard: ``Statistical inference means drawing conclusions based on data''.

De mani\`ere g\'en\'erale, il y a l'id\'ee que les donn\'ees, ce qu'on mesure, peut \^etre pr\'edit par l\'etat du monde. Oui, il y a bien un lien de causali\'e implicite et la mesure arrive toujours en bout de cha\^ine.

Dans l'exemple pr\'ec\'edent, l'\'etat du monde c'est $(v,x)$ la vraie vitesse et la vraie distance. On a $y=v+\epsilon$ mais par bonheur il existe une relation entre $x$ et $v$, 
$p(x,v)=\delta(x-tv)p(v)$. Autrement dit le param\`etre de nuisance $v$ est facilement \'elimin\'e. Plus fondamentalement, les donn\'ees apportent une information sur l'\'etat du monde qui permet de mettre \`a jour sa connaissance.

### 2018-11-16 papers/composite_inference/draft.tex
Modifs 16/11/18

Frequentist and Bayesian inference are based on the concept of likelihood, which relies on the existence of a statistical data-generating model that is both experimentally plausible and computationally tractable. Because this is challenging for complex data, other paradigms are commonly used in applied science: deliberately misspecified generative models, as in quasi-likelihood methods \cite{White-82,Walker-13} or na\"ive Bayes \cite{Ng-01}; minimum description length \cite{Grunwald-07}; and, last but not least, discriminative models\footnote{A {\em discriminative} model describes the conditional distribution of an unobserved variable of interest given an observable variable, in contrast with a {\em generative} model, which describes the conditional distribution of an observable variable given an unobserved variable.}, which currently dominate the field of artificial intelligence and typically require heavy supervised learning -- these include classical techniques such as maximum entropy classifiers \cite{BergerA-96}, support vector machines \cite{Vapnik-00}, Gaussian processes \cite{Rasmussen-06}, as well as most deep learning techniques \cite{Lecun-15,Goodfellow-16} with the exception of deep belief networks \cite{Hinton-06,Fischer-14}.

A strong limitation of discriminative models is that they are not suitable for unsupervised learning or on-the-fly parameter estimation because they treat the data and the model parameters as {\em marginally independent}, meaning that the data conveys no information about the parameters unless the variable of interest is observed. This is illustrated in Figure~\ref{fig:graph_comparison} by the respective directed graph representations of generative and discriminative models. For the same reason, supervised learning in discriminative models is less precise than in comparable generative models, hence less effective in small training sets \cite{Ng-01}. Overall, pure discriminative models are of little use outside the context of big labeled data.

@@

>>>>

Classical frequentist and Bayesian inference paradigms rest upon the existence of a probabilistic data-generating model that is both empirically valid and computationally tractable. Because this is challenging for complex data, other inference models are commonly used in applied science: deliberately misspecified generative models, as in quasi-likelihood methods \cite{White-82,Walker-13} or na\"ive Bayes \cite{Ng-01}; data compression models as in minimum description length \cite{Grunwald-07}; and discriminative models\footnote{A {\em discriminative} model is a parametric family that describes the distribution of a variable of interest conditional on data, in contrast with a {\em generative} model, which describes the conditional distribution of data given the variable of interest.}, which currently dominate the field of artificial intelligence and typically require heavy supervised learning -- these include classical techniques such as maximum entropy classifiers \cite{BergerA-96}, support vector machines \cite{Vapnik-00}, Gaussian processes \cite{Rasmussen-06}, as well as most of deep learning \cite{Lecun-15,Goodfellow-16} (with the exception of deep belief networks \cite{Hinton-06,Fischer-14}).

A strong limitation of discriminative models is that they are not suitable for unsupervised learning or on-the-fly parameter estimation because they treat the data and the model parameters as {\em marginally independent}, meaning that the data conveys no information about the parameters unless the variable of interest is observed. This is illustrated in Figure~\ref{fig:graph_comparison} by the respective directed graph representations of generative and discriminative models. For the same reason, supervised learning in a discriminative model is less precise than in a generative model spanning the same family of posteriors, hence it is less effective in small training sets \cite{Ng-01}. Overall, pure discriminative models are of little use outside the context of big labeled data.

@@

The max entropy training turns out to be unsupervised since the moment
constraints are determined by the 1st learning phase. We have to
interpret that!!!
